---
title: 'Capital Market Activities'
author: 'Santa Rautenbach^[E-mail: Santa.Rautenbach@resbank.co.za. Tel: +27 12 313 3688.] and Wandile Ndwandwe^[E-mail: Wandile.Ndwandwe@resbank.co.za. Tel: +27 12 399 6458.]'
date: "September 2020"
fontsize: 11pt
always_allow_html: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{array}
- \usepackage{longtable}
- \usepackage{pdflscape}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{tabu}
- \usepackage{xcolor}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{tabularx}
- \usepackage{wallpaper}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhf{}
- \fancyhead[C,CO]{\thepage}
- \fancyfoot[L,LO]{Economic Statistics Department \\ Capital Market and Flow of Funds Unit}
- \fancyfoot[R,RO]{Publication Date | 16 October 2020 \\ File Reference | 17/8/1/2}
- \renewcommand{\headrulewidth}{2pt}
- \renewcommand{\footrulewidth}{1pt}
output:
  html_document: default
  pdf_document: default

---

\fancypagestyle{firststyle}{
  \fancyhf{}
  \fancyfoot[L,LO]{Economic Statistics Department \\ Capital Market and Flow of Funds Unit}
  \fancyfoot[R,RO]{Publication Date | 16 October 2020 \\ File Reference | 17/8/1/2}}

\thispagestyle{firststyle}

\renewcommand{\headrulewidth}{0pt}

\ThisCenterWallPaper{0.95}{test.png}
\setlength{\wpXoffset}{-0.1cm}
\setlength{\wpYoffset}{10.9cm}

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Sys.setenv(TZ="Africa/Johannesburg")

library(tidyverse)
Derivatives <- read_csv("~/R/Derivatives.csv")

names(Derivatives)[3] <- "Price"

library(lubridate)
library(pander)
library(htmltools)
library(knitr)
library(rmarkdown)
library(stats)
library(graphics)
library(ggplot2)
library(kableExtra)
library(flextable)
library(dygraphs)
library(plotly)
library(ggthemes)





my_aesthetics <- 
  theme_bw() +
   theme(legend.position = "bottom",
        legend.text=element_text(size=13),
          legend.title = element_blank(),
          axis.line = element_blank(),
          panel.border = element_rect(colour = "black", fill = "NA", size = 2),
          panel.background = element_blank(),
          axis.ticks = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(), 
          panel.grid.major.x = element_line("black"),
          panel.grid.minor.x = element_blank(),
          axis.text = element_text(colour="black"),
          plot.title = element_text(face = "bold"),
        plot.caption = element_text(hjust = 0, size = 13), 
        text = element_text(size = 13,
                              family = "serif",
                              colour = "black"))
  
```



## Domestic agricultural commodity prices displayed a mixed trend.

<br>

* **White maize, yellow maize and sunflower seed prices increased**
* **Meanwhile, wheat prices continued trending downward on the back of record harvests in 2020**

<br>

```{r Derivatives, fig.align='center', echo=FALSE, fig.width = 10, fig.height = 4.5, warning=FALSE}


ggplot(data=Derivatives) +
  geom_line(aes(x = Date, y = Price, color = Commodity), size=1.2) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 0, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust=0)) +
  ggtitle("Monthly derivatives",subtitle = "Rand per ton") +
  labs(caption = "Source: JSE")
```   

## Commodity derivatives

Monthly average prices of domestic agricultural commodities increased on average by 5.2% in September 2020. Wheat prices were the exception which traded lower by 11.9%, while white and yellow maize prices increased by 11.8% an 8.7% respectively. Despite large supplies commodity prices remain at higher levels as growing global demand, primarily for yellow maize, the weaker domestic currency and generally higher global maize prices supported the increase in domestic maize prices. Wheat prices continued lower notwithstanding the lifting of the wheat import tariff as optimism reigns not only on global wheat production levels but also an expectation of the largest domestic wheat harvest in a decade on the back of increased area plantings and favourable rains in the Western Cape. The Agbiz/IDC Agribusiness Confidence Index rebounded from 39 points in the second quarter of 2020 to 51 points in the third quarter, indicating a marginal optimism about business conditions in South Africa.  
