---
title: 'September 2020 QB Review: Monetary developments, interest rates and financial markets'
author: 'Wandile Ndwandwe^[E-mail: Wandile.Ndwandwe@resbank.co.za. Tel: +27 12 399
  6458.]'
date: "12/10/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(plotly)

Yieldcurves <- read.csv("~/R/SeptQB.csv", header=T)
View(Yieldcurves)


```

## Yield curves

The level of the yield curve, which initially declined notably and then flattened somewhat from 23 March 2020 to 3 June, again moved higher and steepened to early July amid renewed risk aversion, in particular beyond the five-year maturity range. Thereafter, the level of the yield curve moved slightly lower once more to 11 September 2020, while maintaining a steep slope. The short end reflected the lower repo rate and increased money market liquidity, whereas the 7–10-year maturity range discounted higher risk premiums and increased government borrowing. The yield gap, measured as the difference between yields at the extreme long and short-ends of the curve, narrowed from 803 basis points on 23 March 2020 to 701 basis points on 3 June before widening again to 800 basis points on 11 September.

```{r Yieldcurves, fig.align="center", echo=FALSE, fig.width = 10, fig.height = 4.5, warning=FALSE}
#creating yield curve plots
ggplot(data=Yieldcurves, aes(x = Period, y = Percent, color=YC)) +
  geom_line(size=1.2) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 0, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust=0)) +
  ggtitle("Yield curves", subtitle = "Unexpired maturity in years") +
  labs(caption = "Sources: IRESS and JSE")
ggplotly()
```